---
title: "R_Week_4_Project_1_607_Daniel_Thonn"
output: html_document
course: R-607 CUNY MSDA program

---

## Assignment Homework-4_Project_1-607: Exploratory Data Analysis

Summary of Assignment
This assignment involves data manipulation and analyis to create a summary sheet for chess tournament data.

This Assignment requires the following:

1). R-Studio

The following R-packages are used:
1.stringr

Steps to reproduce:
1). run the R-Studio file: R_607_Week_4c_Hmk_Text_Manipulation_Daniel_Thonn.Rmd



Setting up and Preparing the Environment

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE}

#install.packages("stringr")
library(stringr)
```


Load the character names data into an R dataframe


```{r, echo=TRUE}
#Load data from a text file
data_1 <- readLines("C:/mysqldata/tournamentinfo.txt") 
head(data_1)


#Obtain only columns with pipe symbol in rows (remove horizontal lines)
data_3 <- data_1[grep(("\\|+"),data_1)]

head(data_3)

# Create a data vector to store manipulated data
data_4 <- 1
list(data_4)

#Combine the rows with names plus each row after

for (i in 3:(length(data_3)-1))
{
  if (!i %% 2){
    next
  }
  print(data_3[i])
  data_4[i] <- paste(data_3[i],data_3[i+1])
}

list(data_4)
data_4b = unlist(data_4)

# Clean data and take only data with pipe symbols
data_5b = data_4b[grep(("\\|+"),data_4b)]
data_5b

#Retain only columns that have a pipe (|)
data_5 <- data_4[grep(("\\|+"),data_4)]
head(data_5)

#Remove trailing extra spaces \\s+
data_6 <- gsub("\\s+"," ",data_5) 
head(data_6)

data_6b <- unlist(data_6)

#Build columns of required data and check each column

playername <- str_trim(unlist(str_extract_all(data_6b, "([[:alpha:]., ]-?){7,}")))
list(playername)


playerstate <- str_trim(unlist(str_extract_all(data_5b, "\\  [[:upper:]]{2}")))
list(playerstate)


playerpoints<-str_trim(unlist(str_extract_all(data_6b, "[0-9]\\.[0-9]-?")))
list(playerpoints)

playerprerating <- as.numeric(gsub("\\: ","",unlist(str_extract_all(data_5b, "\\: [[:digit:] ]{4}"))))
playerprerating 

oppprerating1=unlist(str_extract_all(data_5b, "\\|[[:upper:]]  [[:digit:] ]{1,}"))
head(oppprerating1)

oppprerating2 <- as.numeric(unlist(str_extract_all(oppprerating1,"[[:digit:]]{1,2}")))
head(oppprerating2)

# obtain the opponent pre-rating by indexing the player pre-rating by the opponents played
oppprerating3 <- playerprerating[oppprerating2]
head(oppprerating3)
list(oppprerating3)


# Create a matrix of opp-preratings to obtain average
oppprerating4 <- matrix(oppprerating3,7,64)

#Take transpose of matrix to obtain 64x7 format that fits the data output needed
oppprerating5 <- t(oppprerating4)
head(oppprerating5)

#Take the mean of the opponent ratings by player
avg_oppprerating <- as.integer(rowMeans(oppprerating5))
head(avg_oppprerating)

#Create a data frame with the fields required
chess_summary <- data.frame(playername,playerstate,playerpoints,playerprerating,avg_oppprerating)
list(chess_summary)

#Write the chess summary output to a csv file
write.csv(chess_summary, "C:/mysqldata/chess_summary.csv")


```



#END
